Depletion capabilities in the OpenMC Monte Carlo particle transport code

### 2020-06-23 manuscript.tex
Add discussion of SFR results

@@

Six coupled transport-depletion simulations were carried out with OpenMC to test
a variety of combinations of depletion chains and methods for handling FPY
energy dependence:
\begin{table}

\end{table}
\begin{enumerate}
  \item Full depletion chain, thermal FPYs
  \item Full depletion chain, interpolated FPYs
  \item Reduced depletion chain, thermal FPYs
  \item CASL depletion chain, thermal FPYs
  \item CASL depletion chain, FPYs based on cutoff method
  \item CASL depletion chain, interpolated FPYs
\end{enumerate}
Because Serpent doesn't have an option for changing the depletion chain or the
method for handling FPY energy dependence, a single coupled transport-depletion
simulation with Serpent was carried out. For all simulations, 20 inactive
batches, 200 active batches, and \num{3.6e6} particles per batch were used,
which resulted in an uncertainty on the effective multiplication factor of about
3--4 pcm.

@@

\Cref{fig:pwr-fp-casl-average} shows the relative difference in predicted
fission product concentrations between OpenMC and Serpent when using the CASL
depletion chain and interpolated FPYs. As in \cref{fig:pwr-fp-full-average},
almost all fission products are predicted to well within 1\% at 50 MWd/kg.
However, in this case there is one clear outlier: $^{85}$Kr. Upon further
investigation, it was determined that the root cause for this difference is the
fact that the CASL depletion chain uses a cumulative yield for $^{85}$Kr, and
the cumulative yield of $^{85}$Kr is known to be inconsistent with the
independent yields of the $A=85$ mass chain~\cite{pigni2015nds}. Serpent relies
solely on the independent yield in its depletion solver (as does the full
depletion chain). Therefore, there is nothing inherently wrong with the CASL
depletion chain or OpenMC; it is simply a matter of the underlying nuclear data
being inconsistent. The 5\% underprediction of $^{85}$Kr is also likely
responsible for the slightly higher difference in $k_\text{eff}$ for the CASL
depletion chain shown in \cref{fig:pwr-keff}.
\begin{figure}[H]
  \centering
  \includegraphics[width=4in]{figures/pwr/pwr_fp_casl_average.pdf}
  \caption{Relative difference in fission product concetrations between OpenMC
  and Serpent for the PWR pincell model at 5, 12.5, 30, and 50 MWd/kg using the
  CASL depletion chain and interpolated FPYs.}
  \label{fig:pwr-fp-casl-average}
\end{figure}

@@

@@

>>>>

TODO: Discuss missing FPY~\cite{fensin2015pne}.

@@

Six coupled transport-depletion simulations were carried out with OpenMC to test
a variety of combinations of depletion chains and methods for handling FPY
energy dependence:
\begin{enumerate}
  \item Full depletion chain, thermal FPYs
  \item Full depletion chain, interpolated FPYs
  \item Reduced depletion chain, thermal FPYs
  \item CASL depletion chain, thermal FPYs
  \item CASL depletion chain, FPYs based on cutoff method
  \item CASL depletion chain, interpolated FPYs
\end{enumerate}
Because Serpent doesn't have an option for changing the depletion chain or the
method for handling FPY energy dependence, a single coupled transport-depletion
simulation with Serpent was carried out. For all simulations, 20 inactive
batches, 200 active batches, and \num{4.0e6} particles per batch were used,
which resulted in an uncertainty on the effective multiplication factor of about
3--4 pcm.

@@

\Cref{fig:pwr-fp-casl-average} shows the relative difference in predicted
fission product concentrations between OpenMC and Serpent when using the CASL
depletion chain and interpolated FPYs. As in \cref{fig:pwr-fp-full-average},
almost all fission products are predicted to well within 1\% at 50 MWd/kg.
However, in this case there is one clear outlier: $^{85}$Kr. Upon further
investigation, it was determined that the root cause for this difference is the
fact that the CASL depletion chain uses a cumulative yield for $^{85}$Kr, and
the cumulative yield of $^{85}$Kr is known to be inconsistent with the
independent yields of the $A=85$ mass chain~\citep{pigni2015nds}. Serpent relies
solely on the independent yield in its depletion solver (as does the full
depletion chain). Therefore, there is nothing inherently wrong with the CASL
depletion chain or OpenMC; it is simply a matter of the underlying nuclear data
being inconsistent. The 5\% underprediction of $^{85}$Kr is also likely
responsible for the slightly higher difference in $k_\text{eff}$ for the CASL
depletion chain shown in \cref{fig:pwr-keff}.
\begin{figure}[H]
  \centering
  \includegraphics[width=4in]{figures/pwr/pwr_fp_casl_average.pdf}
  \caption{Relative difference in fission product concetrations between OpenMC
  and Serpent for the PWR pincell model at 5, 12.5, 30, and 50 MWd/kg using the
  CASL depletion chain and interpolated FPYs.}
  \label{fig:pwr-fp-casl-average}
\end{figure}

@@

The SFR assembly model is based on an OECD/NEA benchmark for SFR
analysis~\citep{nsc2015}. Namely, a 2D model of a driver assembly for the 1000
MWth metallic core (MET-1000) was used. The fuel composition corresponded to the
lowest axial region (the 17.16 cm column in Table 2.23 from ~\citep{nsc2015}).
All dimensions and material compositions can be found in the original OECD/NEA
report. The assembly model extends to positive/negative infinity in the $z$
direction and has periodic boundary conditions on the outer hexagonal surfaces.
A linear heating rate of \SI{23.3}{\kilo\watt\per\meter} per
pin~\citep{cahalan2007anl} was used for power normalization. The assembly was
depleted for 360 days in 20 day timesteps. As in the PWR pincell case, the
predictor method was used in both codes for time integration.

Unlike the PWR pincell model, which only contains materials at a single
temperature, the SFR assembly model uses a higher temperature in the fuel than
in other materials. In OpenMC, the temperature-dependence of neutron cross
sections was handled by interpolating data between neighboring temperatures.
Serpent uses an explicit temperature treatment known as target motion
sampling~\cite{viitanen2012nse}. However, we note that for this problem, the
temperature dependence of unresolved resonance probability tables becomes
important because of the fast flux spectrum, but the temperature treatment in
Serpent 2.1.31 does not handle probability tables.

Four coupled transport-depletion simulations were carried out with OpenMC for
the SFR assembly model:
\begin{enumerate}
  \item Full depletion chain, interpolated FPYs, probability tables on
  \item CASL depletion chain, interpolated FPYs, probability tables on
  \item Full depletion chain, interpolated FPYs, probability tables off
  \item CASL depletion chain, interpolated FPYs, probability tables off
\end{enumerate}
The results for the PWR pincell model indicated that the reduced depletion chain
results in identical behavior to the full depletion chain, so we have not
included a case for the SFR assembly using the reduced depletion chain. Two
coupled transport-depletion simulation with Serpent was carried out, one with
probabilty tables turned on and one with them turned off. For all simulations,
20 inactive batches, 200 active batches, and \num{4.0e6} particles per batch
were used, which resulted in an uncertainty on the effective multiplication
factor of about 3--4 pcm.

% Also, we
% have only included a single case with thermal FPYs to compare with the use of
% interpolated FPYs using the same depletion chain.

\Cref{fig:sfr-keff} shows the difference between $k_\text{eff}$ over time
calculated by OpenMC and Serpent. The blue points correspond to results using
the full depletion chain, and the green points correspond to results using the
CASL chain. For each chain, simulations with probability tables on are denoted
by circles and simulations without probability tables are denoted by squares.
Unlike in the PWR pincell case, there is a statistically significant difference
in $k_\text{eff}$ at zero days. Let us first consider the case with probability
tables turned off. In this case, $k_\text{eff}$ calculated by OpenMC is about 20
pcm higher than that of Serpent. When using the full depletion chain, there is
no change in the difference between OpenMC and Serpent. When using the CASL
depletion chain, the difference increases by about 20--30 pcm, similar to the
behavior observed in \cref{fig:pwr-keff}. With probability tables turned on,
however, the situation is quite different. At zero days, $k_\text{eff}$
calculated by OpenMC is 20 pcm lower than the value calculated by Serpent. As
the burnup increases, there is an increasing positive bias between OpenMC and
Serpent. To understand the cause of this difference, it is instructive to look
at the differences in isotopic compositions.
\begin{figure}[H]
  \centering
  \includegraphics[width=4in]{figures/sfr_keff.pdf}
  \caption{Difference in the effective multiplication factor, $k_\text{eff}$,
  between OpenMC and Serpent for the SFR assembly model. While not shown, the one
  standard deviation uncertainty in the difference is about 3.5 pcm.}
  \label{fig:sfr-keff}
\end{figure}

\Cref{fig:sfr-actinides} shows the relative difference in predicted actinide
concentrations between OpenMC and Serpent with probability tables turned on. For
both the full and CASL depletion chains, the relative differences are all within
0.2\%. While this may seem close, the observed difference in $^{238}$U and
$^{239}$Pu concentrations merits special mention. In this model, the fuel
material is at \SI{807}{\kelvin}, but the endf71x library contains cross
sections processed at \SI{600}{\kelvin} and \SI{900}{\kelvin}. Thus, Serpent
ends up using probability tables at \SI{600}{\kelvin} rather than
\SI{807}{\kelvin}, which results in less parasitic absorption in $^{238}$U and
consequently less production of $^{239}$Pu (hence a lower $k_\text{eff}$ than
OpenMC with increasing burnup). Although OpenMC also does not have access to
probability tables at the exact fuel temperature, it statistically interpolates
between the probability tables at \SI{600}{\kelvin} and \SI{900}{\kelvin}.
\begin{figure}[H]
  \centering
  \begin{subfigure}[t]{0.45\textwidth}
    \includegraphics[width=\textwidth]{figures/sfr/sfr_actinides_full.pdf}
    \caption{Full depletion chain}
    \label{fig:sfr-actinides-full}
  \end{subfigure}
  \hspace{0.05\textwidth}
  \begin{subfigure}[t]{0.45\textwidth}
    \includegraphics[width=\textwidth]{figures/sfr/sfr_actinides_casl.pdf}
    \caption{CASL depletion chain}
    \label{fig:sfr-actinides-casl}
  \end{subfigure}
  \caption{Relative difference in actinide concetrations between OpenMC and
  Serpent for the SFR assembly model at 80, 180, 260, and 360 days with
  probability tables on.}
  \label{fig:sfr-actinides}
\end{figure}

When probability tables are turned off, very close agreement is observed between
the predicted actinide concentrations in OpenMC and Serpent, as shown in
\cref{fig:sfr-actinides-nop}. In this case, the actinide concetrations agree to
within 0.01\% at all times during depletion for both the full and CASL depletion
chains.
\begin{figure}[H]
  \centering
  \begin{subfigure}[t]{0.45\textwidth}
    \includegraphics[width=\textwidth]{figures/sfr/sfr_actinides_full_nop.pdf}
    \caption{Full depletion chain}
    \label{fig:sfr-actinides-full-nop}
  \end{subfigure}
  \hspace{0.05\textwidth}
  \begin{subfigure}[t]{0.45\textwidth}
    \includegraphics[width=\textwidth]{figures/sfr/sfr_actinides_casl_nop.pdf}
    \caption{CASL depletion chain}
    \label{fig:sfr-actinides-casl-nop}
  \end{subfigure}
  \caption{Relative difference in actinide concetrations between OpenMC and
  Serpent for the SFR assembly model at 80, 180, 260, and 360 days with
  probability tables off.}
  \label{fig:sfr-actinides-nop}
\end{figure}

\Cref{fig:sfr-fp} shows the relative difference in fission product
concentrations between OpenMC and Serpent with probability tables turned on.
Most differences between OpenMC and Serpent are below 1\%. As with the PWR
pincell model, OpenMC tends to produce a lower concentration of $^{109}$Ag most
likely because of differences in the treatment of FPY energy dependence. The
only noticeable difference between the full and CASL depletion chains is for
$^{85}$Kr, which is again explained by the use of an inconsistent cumulative
yield in the CASL depletion chain.
\begin{figure}[H]
  \centering
  \begin{subfigure}[t]{0.45\textwidth}
    \includegraphics[width=\textwidth]{figures/sfr/sfr_fp_full_average.pdf}
    \caption{Full depletion chain}
  \end{subfigure}
  \hspace{0.05\textwidth}
  \begin{subfigure}[t]{0.45\textwidth}
    \includegraphics[width=\textwidth]{figures/sfr/sfr_fp_casl_average.pdf}
    \caption{CASL depletion chain}
  \end{subfigure}
  \caption{Relative difference in fission product concetrations between OpenMC
  and Serpent for the SFR assembly model at 80, 180, 260, and 360 days with
  probability tables on.}
  \label{fig:sfr-fp}
\end{figure}

With probability tables turned on, the differences between OpenMC and Serpent in
fission product concentrations are mostly the same as when probability tables
were turned off, as shown in \cref{fig:sfr-fp-nop}. We note that because the
difference in $^{238}$U and $^{239}$Pu was still less than 0.2\% in
\cref{fig:sfr-actinides}, it is expected that the use of probability tables
doesn't have a large effect on the fission product concentrations. Again, most
fission product concentrations agree to within 1\%, with $^{109}$Ag and
$^{85}$Kr being the notable outliers.
\begin{figure}[H]
  \centering
  \begin{subfigure}[t]{0.45\textwidth}
    \includegraphics[width=\textwidth]{figures/sfr/sfr_fp_full_average_nop.pdf}
    \caption{Full depletion chain}
  \end{subfigure}
  \hspace{0.05\textwidth}
  \begin{subfigure}[t]{0.45\textwidth}
    \includegraphics[width=\textwidth]{figures/sfr/sfr_fp_casl_average_nop.pdf}
    \caption{CASL depletion chain}
  \end{subfigure}
  \caption{Relative difference in fission product concetrations between OpenMC
  and Serpent for the SFR assembly model at 80, 180, 260, and 360 days with
  probability tables off.}
  \label{fig:sfr-fp-nop}
\end{figure}

@@

Discuss temperature dependence and probability tables in SFR model.
